Report

### 2023-04-25 manuscript.Rmd
Rebecca's edits

abstract: |
  The product Bayes factor (PBF) can synthesize evidence in favor of an informative hypothesis across heterogeneous replication studies. It is particularly useful when the number of studies is relatively low and conventional assumptions about between-studies heterogeneity are likely violated. The present paper introduces a user-friendly implementation of the PBF in the bain R-package, and demonstrates its use in several tutorial examples. The method was validated in a simulation study that manipulated sample size, number of replication samples, and reliability. PBF had a higher overall accuracy when benchmarked against other evidence synthesis methods, including random-effects meta-analysis (RMA). This was primarily due to PBF’s greater sensitivity in detecting a true effect. However, PBF had relatively lower specificity. The PBF showed increasing sensitivity and specificity with increasing sample size. With an increasing number of samples, lower sensitivity was traded for greater specificity. Although PBF's overall performance was less susceptible to reliability than the other algorithms, this masked a trade-off between reliability and specificity. PBF thus appears to be a promising method for meta-analysis of heterogeneous conceptual replication studies. Nonetheless, users should be aware of its lower specificity, and the fact that it takes a Bayesian approach to inference and addresses a qualitatively different research question than other evidence synthesis methods.

@@

```{r setup, include = FALSE}
run_everything <- TRUE
library("papaja")
library(tidySEM)
library(kableExtra)
library(pwr)
library(bain)
r_refs("r-references.bib")
source("UI function/pbf.R")
out <- readRDS("out.RData")
```

@@

Recent years have seen a crisis of confidence over the reliability of published results in psychology, and science more broadly [@brembs2018prestigious].
Replication research has emerged as one potential way to address this crisis and derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021]. 
In step with this interest in replication research,
research synthesis methods have become increasingly popular.
These methods aggregate research findings,
and thus enable drawing overarching conclusions across multiple (replication) studies.
In this paper, we introduce a new research synthesis method that uses Bayesian evidence synthesis to aggregate the evidence in favor of an informative hypothesis, quantified by the Bayes factor,
across multiple studies.
This method has the potential to provide a more comprehensive and accurate picture of the state of the literature, and to identify areas of consensus and disagreement among studies.
We describe the method in detail and demonstrate its application through a tutorial example analysis.

@@

A crucial shortcoming of existing research synthesis methods is that the aforementioned assumptions
may not be tenable when meta-analyzing studies that investigate the same informative hypothesis,
but are otherwise very heterogeneous.
The situation may arise where each study is uniquely identified by a combination of linearly dependent moderators.
In this case, it is no longer possible to synthesize *effect sizes* while accounting for heterogeneity using statistical methods.
<!-- even machine learning based methods cannot identify relevant moderators. -->
It is still possible, however, to quantify the support these studs provide for the underlying informative hypothesis.
To this end, Bayesian evidence synthesis (BES) aggregates the evidence in favor of an informative hypothesis $H_i$ across studies,
without imposing assumptions about heterogeneity [@kuiperCombiningStatisticalEvidence2013].

The amount of evidence for a hypothesis can be expressed as a Bayes factor, or BF.
The BF can be interpreted as the ratio of evidence in favor of one hypothesis $H_1$ relative to another hypothesis $H_2$, so $BF_{12} = \frac{H_1}{H_2}$.
Within the scope of this paper,
all Bayes factors are the ratio of evidence in favor of an informative hypothesis $H_i$ relative to its complement $H_{!i}$, $BF_{c} = \frac{H_i}{H_!i}$. 
The subscript $_!$ represents the negation operator;
in other words, $H_{!i}$ means "not $H_i$".
Consequently, $BF_c$ represents the ratio of evidence in favor of $H_i$ divided by evidence against it.
A value of $BF_c = 10$ means that the data provide ten times more support in favor of the hypothesis than against it.

@@

Although meta-analysis and BES are both research synthesis methods, they answer different research questions.
Meta-analysis estimates the point estimate or distribution of a population effect size. 
It pools estimates of this effect size across multiple studies to obtain an overall estimate of the effect size.
It thus answers questions like:
Given certain assumptions about between-studies heterogeneity, what is the average population effect size?
BES, on the other hand, aggregates evidence in favor of an informative hypothesis across multiple studies.
It thus answers the question: Do all these studies support the hypothesis of interest?
Both methods are appropriate for different research questions, and provide complementary information.

@@

<!-- Rebecca: -->
<!-- ms meer over het model vertellen. -->
<!-- Is het voor de lezer duidelijk waar ineens deze groepen vandaan komen? Zijn dat je aantal studies die je gaat samennemen? -->
<!-- Dus bijv: -->
<!-- In de populatie zijn er twee variabelen, die cooreleren met de waarde rho. We hebben k studies die elk een sample van n hebben getrokken/verzameld. Dan vertelen wat de verschillende algrotithms doen met de verschillende samples (meeste als niet alle: In elke sample wordt de sample correlatie bepaald; dan estimates samenemen en null hypo test of in each sample bewijs voor hypo en dat samennemen). Dan iets over de verschillende waardes die je bekijkt en dat elke combi een simulatie conditie is. En dan iets als: Dit doen we 1000 keer per set van condities.  -->
<!-- En dan nog iets over reliability ook. -->
The present simulation study set out to validate the PBF algorithm and benchmark it against other evidence synthesis methods.
We simulated a scenario where an informative hypothesis about a correlation between two variables was measured across several independent samples,
and the resulting evidence was synthesized across samples using multiple methods.
<!-- Rebecca: -->
<!-- waar is > 3 op gebaseerd? -->
<!-- Kass&Raftery? Als ja, bedenk wel dat dat om equaility restricties gaat (bij mijn weten). Je kan het natuurlijk gebruiken, maar plaats dan wel de kanttekening, maar is een uitgangspunt natuurlijk.  -->
<!-- Of vertel over betting odds... -->
<!-- Maar het blijft een dichitome beslissing zo.... -->
<!-- Laat je het gemiddelde PBF met range ofzo zien.... zo niet, dan zou dat toevoegen. Dan zie je meer dan een dichitome beslissing. -->
<!-- En ms kan je ook ene keer 1 en 2 gebruiken, gewoon om te zien wat er gebeurt? -->
<!-- Rebecca: -->
<!-- Now, you will say that selecting > .1 with rho .1 is false, but it is not false. It is on the border, so Hi is true.  -->
<!-- In practice it may be true half of the time, because of sampling, but still. -->
<!-- This should be reflected by BF is approx 1. -->
<!-- In your simulation, you use PBF > 3, so this may be hard to reach when the truth is on the border, but that is due to the subjective choice of 3... -->
<!-- So, I already thought that using a smaller population rho value would be of added value, but now I believe you should really include this. -->
<!-- I do see that this would end up in multiple results (once using .1 and 0.2, once (say) 0 or 0.05 and .2; but also when using other cut-off values for PBF). -->
The informative hypothesis was kept constant at $H_i: \rho > .1$.
To examine the performance of the different evidence synthesis methods in a range of scenarios,
several design factors were manipulated.
First was the presence or absence of a true population effect.
A present effect was defined as $\rho = .2$;
a null effect was defined as $\rho = .1$.
The second design factor was the number of observations per sample $n \in (50, 200, 500, 800)$.
These values were chosen because they correspond to a statistical
<!-- pwr.r.test(c(50, 200, 500, 800), .1) -->
power to reject a false null hypothesis of $\beta \in (.10, .30, .60, .80)$ power, respectively,
assuming $\alpha = .05$ and a known effect size of $\rho = .1$.
Third, we manipulated the number of independent samples (or: replication studies), $k \in (2, 3, 10)$.
Fourth, the reliability of the two correlated variables was varied between $\alpha \in (0.6, 0.8, 1.0)$,
where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is
assumed when analyzing correlations between observed items or scale scores.
For all unique combinations of these design factors,
the simulation was repeated 1000 times.

@@

As a benchmark for comparison, we included several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
The first was *vote counting*: counting the number of significant effects.
Although this method is still in use for aggregating conceptual replications,
it is considered bad practice.
Three disadvantages are that vote counting disregards sample size,
reduces statistical power,
and does not quantify the strength of the evidence [@hedgesVotecountingMethodsResearch1980].
Our vote counting algorithm summed the number of one-sided z-tests of a null hypothesis corresponding to the informative hypothesis,
so $H_0: \rho = .1$ and $H_a: \rho > .1$, which corresponds to $H_i$.
The decision criterion was that the hypothesis was supported in the majority of samples.
Thus, for example, if $H_a$ was supported in three out of five samples, our vote counting algorithm would find overall support for $H_a$ (and, by extension, $H_i$).

@@

For each algorithm, inferential decisions made using the criteria described above were compared to the population status of the hypothesis (true or false).
The resulting confusion matrix gives the number of decisions that were true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN).
These quantities were summarized as sensitivity, $\frac{TP}{TP + FN}$,
the ability to detect an effect given that it was indeed true in the population,
<!-- i think sensitivity is the ability to detect an effect, given that the effect is truly there. I found that its formula is usually presented as TP/(TP+FN)-->
<!--CJ: This is all the same, see https://en.wikipedia.org/wiki/Sensitivity_and_specificity -->
and specificity, $\frac{TN}{TN+FP}$,
the ability to correctly conclude that the informative hypothesis is not supported,
given that it was indeed false in the population. <!-- and specificity is the ability to not acknowledge an effect, given that there indeed is none-->
The overall performance was captured by the accuracy, which represents the total proportion of correct (true positive and true negative) decisions, $\frac{TP + TN}{TP+TN+FP+FN}$.

@@

```{r tabconf, echo = F, message=F}
tab <- read.csv("confusion.csv", stringsAsFactors = FALSE)
names(tab)[1] <- "Metric"
tab$Metric <- gsub("^.+_", "", tab$Metric)
tab <- tab[!tab$Metric == "lr", !grepl("iu", names(tab))]
names(tab) <- gsub("_ic", "", names(tab))
# names(tab)[match(names(rename), names(tab))] <- rename
tab <- tab[tab$Metric %in% c("sensitivity", "specificity", "accuracy"),]
rownames(tab) <- NULL
knitr::kable(tab, digits = 2, caption = "Marginal confusion matrix metrics.")
```
First, we examined overall model performance across conditions.
PBF had a higher overall accuracy than other algorithms
followed by IPD, then RMA and finally VC,
see Table \@ref(tab:tabconf).
This higher accuracy was primarily driven by PBF's greater sensitivity to detect a true effect compared to other algorithms.
However, PBF had a lower specificity compared to all other algorithms.
This suggests that the PBF trades a loss of specificity for increased sensitivity.

@@

The figure indicates that, for PBF at higher levels of $k$,
lower sensitivity was traded for greater specificity.
The other algorithms did not show this pattern, as their specificity was at a ceiling.
This difference in pattern of effects explains why number of samples had a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
Only VC showed decreasing sensitivity with an increasing number of groups;
this is because the probability of obtaining false negatives increases with the number of groups [as also noted by @hedgesVotecountingMethodsResearch1980].

@@

A third limitation is that the evidence synthesis methods compared here represent different approaches to inference and answer different research questions.
<!-- Their inferential properties are thus not directly comparable. -->
<!-- PBF assesses whether an informative hypothesis is supported across all replication studies; -->
<!-- RMA estimates an average effect size from summary statistics; -->
<!-- IPD estimates an effect size from individual data, accounting for clustering within studies; -->
<!-- and VC counts the number of times a hypothesis was supported. -->
Since each of these methods is optimized for a different purpose,
the present study should not be considered as a comprehensive assessment of their strengths and weaknesses.
<!-- For example, PBF takes a Bayesian approach to inference, and a PBF of 10 in our study means that the (subjective) probability that our informative hypothesis is true is ten times larger than the probability that it is not true. -->
<!-- By contrast, IPD takes a frequentist approach to inference, -->
<!-- and a p-value of $p = .02$ means that there is a 2% probability of observing an effect at least as large as the one found, if the null-hypothesis were true. -->
We nonetheless compare them because of their similar usage in evidence synthesis.
It is up to individual researchers to choose an appropriate method,
guided by the research question and the available information.
For instance, when raw data is unavailable, IPD cannot be used,
and when effect sizes are not available, only VC can be used.

Aside from the aforementioned fact that the PBF answers a different research question than the other algorithms,
it is worth noting limitations of the interpretation of the PBF.
The PBF renders support for one specific informative hypothesis versus its complement.
If the informative hypothesis is supported, this does not necessarily mean that it is also true.
Consider the hypothetical example that a test of the informative hypothesis that the earth is flat received support with $BF = 3.01$.
Although the data support this hypothesis over its complement, the hypothesis is clearly wrong (the earth is spherical).
If we would have tested another hypothesis, e.g., the earth is shaped like an American football, it would have received much more support, e.g. $BF = 1000$, even though it is also wrong.
A high Bayes factor thus does not mean that the hypothesis is true.
Conversely, a low Bayes factor merely indicates that the informative hypothesis is not supported,
and does not provide information about the true state of affairs.

<!-- In sum, this simulation study suggest that PBF may be a good choice when sensitivity is of utmost importance, but researchers should be mindful of the trade-off between specificity and sensitivity. -->
<!-- The results also highlight the importance of considering sample size, number of samples, -->
<!-- and reliability when designing studies and conducting meta-analyses. Future research should examine the performance of additional algorithms and the impact of other design factors on algorithm performance. -->
<!-- The results indicated that, compared to the other algorithms, -->
<!-- PBF had a higher overall accuracy, -->
<!-- which could be attributed to PBF's relatively greater sensitivity to detect a true effect. -->
<!-- The greater sensitivity of PBF came at a loss of specificity compared to other algorithms. -->
<!-- Specifically, all other algorithms showed ceiling effects for specificity and low base rates for sensitivity. -->

@@

This tutorial demonstrates how to synthesize evidence for an informative hypothesis across heterogeneous replications using the Product Bayes Factor (PBF). 
We assume that users have installed the free open source statistical programming language R [@R-base].
The R-package `bain` version `0.2.9` or later is required, which can be installed by running `install.packages("bain")` in the R console.
The data used in this tutorial are included in the `bain` package,
and have been simulated based on the data presented in [@vanleeuwen2022morality].
A more detailed description of the datasets is found in [@vanleeuwen2022morality];
additionally, the dataset documentation is accessed by running `?synthetic_us`, `?synthetic_dk` or `?synthetic_nl` in the R console.
Van Leeuwen and colleagues conducted a theory-driven, preregistered study to address the research question whether political orientation and moral dispositions are associated.
Suitable data were collected in three countries: the United states of America, Denmark and the Netherlands.
Each sample contained multiple measures of political orientation and moral dispositions.
In the original publication, the PBF was used to aggregate evidence across scales and countries to obtain an overall measure of support for the central hypothesis.
This tutorial follows the same rationale, but uses only one effect size per sample,
and varies the way this effect size is computed to illustrate the more typical use case where the same informative hypothesis has been studied in different ways in multiple studies.
We will examine the informative hypothesis that self-reported importance of family morality is positively associated with a conservative socio-political orientation.
First, we load the `bain` library and assign the data to three objects with convenient names:

```{r eval = TRUE}
library(bain)
NL <- synthetic_nl
DK <- synthetic_dk
US <- synthetic_us
```

@@

<!-- CJ: Is dit niet teveel herhaling van wat we in de inleiding zeggen?? -->
<!-- In conventional Null-hypothesis significance testing (NHST), as the name implies, data is analysed to generate evidence for or against the null-hypothesis.  -->
<!-- This method evaluates a single hypothesis, the null-hypothesis, and its output is dichotomous in nature: whether a theoretical relationship is probable or improbable to exist given the data.  -->
<!-- However, there are many more hypotheses that can theoretically be tested and researchers may not be interested in the null-hypothesis. They may also not only be interested in whether the hypothesis is probable, but also by how much.  -->
<!-- In the Bayesian framework, the Bayes Factor (BF) is a test statistic that gives the amount of support for any possible desired hypothesis.  -->
<!-- Because the focus of this tutorial is on BES,  -->
First, we briefly introduce the basic use of the `bain()` function, and how to interpret its output.
We must estimate a model suitable for testing our informative hypothesis.
Because both scales consist of multiple items, we can use structural equation modeling (SEM) to perform latent variable regression [see @vanlissaTeacherCornerEvaluating2020]:
<!--EB: For simplicities sake, I for now create simple average scores without taking into account reverse coding. This should likely be improved with factor scores -->
<!-- CJ: Keep the code SUPER simple. No functions etc -->

@@

# Test the hypothesis that the effect size labeled 'beta' is positive
# The argument standardize = TRUE uses the standardized model parameters
bf_nl <- bain(results_nl, hypothesis = "beta > 0", standardize = TRUE)
```

>>>>

abstract: |
  The product Bayes factor (PBF) can synthesize evidence for an informative hypothesis across heterogeneous replication studies. It is particularly useful when the number of studies is relatively low and conventional assumptions about between-studies heterogeneity are likely violated. The present paper introduces a user-friendly implementation of the PBF in the bain R-package, and demonstrates its use in several tutorial examples. The method was validated in a simulation study that manipulated sample size, number of replication samples, and reliability. PBF had a higher overall accuracy when benchmarked against other evidence synthesis methods, including random-effects meta-analysis (RMA). This was primarily due to PBF’s greater sensitivity in detecting a true effect. However, PBF had relatively lower specificity. The PBF showed increasing sensitivity and specificity with increasing sample size. With an increasing number of samples, lower sensitivity was traded for greater specificity. Although PBF's overall performance was less susceptible to reliability than the other algorithms, this masked a trade-off between reliability and specificity. PBF thus appears to be a promising method for meta-analysis of heterogeneous conceptual replication studies. Nonetheless, users should be aware of its lower specificity, and the fact that it takes a Bayesian approach to inference and addresses a qualitatively different research question than other evidence synthesis methods.

@@

```{r setup, include = FALSE}
run_everything <- TRUE
library("papaja")
library(tidySEM)
library(kableExtra)
library(pwr)
library(bain)
r_refs("r-references.bib")
source("UI function/pbf.R")
out <- readRDS("out.RData")
resultaction = "markup"
```

@@

Recent years have seen a crisis of confidence over the reliability of published results in psychology, and science more broadly [@brembs2018prestigious].
Replication research has emerged as one potential way to address this crisis and derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021]. 
In step with this interest in replication research,
research synthesis methods have become increasingly popular.
These methods aggregate research findings,
and thus enable drawing overarching conclusions across multiple (replication) studies.
This paper addresses Bayesian evidence synthesis,
a research synthesis method that aggregates evidence for an informative hypothesis, quantified by the Bayes factor,
across multiple studies.
This method has the potential to provide a more comprehensive and accurate picture of the state of the literature, and to identify areas of consensus and disagreement among studies.
We describe the method in detail,
benchmark its performance against other commonly used research synthesis methods,
and demonstrate its application through a tutorial example analysis.
To facilitate uptake of the method by applied researchers, we introduce an implementation of the method in the `bain` R-package.
This implementation enables the use of Bayesian evidence synthesis for many commonly used statistical analyses in R.

@@

A crucial shortcoming of existing research synthesis methods is that the aforementioned assumptions
may not be tenable when meta-analyzing studies that investigate the same informative hypothesis,
but are otherwise very heterogeneous.
The situation may arise where each study is uniquely identified by a combination of linearly dependent moderators.
In this case, it is no longer possible to synthesize *effect sizes* while accounting for heterogeneity using statistical methods.
<!-- even machine learning based methods cannot identify relevant moderators. -->
It is still possible, however, to quantify the support these studies provide for the underlying informative hypothesis.
To this end, Bayesian evidence synthesis (BES) aggregates the evidence for a theoretical relationship across studies,
without imposing assumptions about heterogeneity [@kuiperCombiningStatisticalEvidence2013].
Although this assumption is not necessary,
for the sake of simplicity we assume that this theoretical relationship is evaluated via informative hypothesis $H_i$ in all studies.

The amount of evidence for a hypothesis can be expressed as a Bayes factor, or BF.
The BF can be interpreted as the ratio of evidence for one hypothesis relative to another hypothesis.
<!-- , so $BF_{12} = \frac{BF_{1u}}{BF_{2u}}$. -->
<!-- Note that $BF_{12}$ is itself a ratio of two Bayes factors with a common denominator, which renders them comparable [see @guApproximatedAdjustedFractional2018]. -->
Within the scope of this paper,
all Bayes factors are the ratio of evidence for an informative hypothesis $H_i$ relative to its complement $H_{!i}$ [see @guApproximatedAdjustedFractional2018].
<!-- This Bayes factor , $BF_{c} = \frac{H_i}{H_!i}$.  -->
The subscript $_!$ represents the negation operator;
in other words, $H_{!i}$ means "not $H_i$".
This Bayes factor, which we will refer to as $BF_c$, represents the ratio of evidence for $H_i$ divided by evidence against it.
A value of $BF_c = 10$ means that the data provide ten times more support for the hypothesis than against it.

@@

Although meta-analysis and BES are both research synthesis methods, they answer different research questions.
Meta-analysis estimates the point estimate or distribution of a population effect size. 
It pools estimates of this effect size across multiple studies to obtain an overall estimate of the effect size.
It thus answers questions like:
Given certain assumptions about between-studies heterogeneity, what is the average population effect size?
BES, on the other hand, aggregates evidence for an informative hypothesis across multiple studies.
It thus answers the question: Do all these studies support the hypothesis of interest?
Both methods are appropriate for different research questions, and provide complementary information.

@@

<!-- Rebecca: -->
<!-- ms meer over het model vertellen. -->
<!-- Is het voor de lezer duidelijk waar ineens deze groepen vandaan komen? Zijn dat je aantal studies die je gaat samennemen? -->
<!-- Dus bijv: -->
<!-- In de populatie zijn er twee variabelen, die cooreleren met de waarde rho. We hebben k studies die elk een sample van n hebben getrokken/verzameld. Dan vertelen wat de verschillende algrotithms doen met de verschillende samples (meeste als niet alle: In elke sample wordt de sample correlatie bepaald; dan estimates samenemen en null hypo test of in each sample bewijs voor hypo en dat samennemen). Dan iets over de verschillende waardes die je bekijkt en dat elke combi een simulatie conditie is. En dan iets als: Dit doen we 1000 keer per set van condities.  -->
<!-- En dan nog iets over reliability ook. -->
The present simulation study set out to validate the PBF algorithm and benchmark it against other evidence synthesis methods.
We simulated a scenario where an informative hypothesis about a correlation between two variables was measured across several independent samples,
and the resulting evidence was synthesized across samples using multiple methods.
<!-- Rebecca: -->
<!-- waar is > 3 op gebaseerd? -->
<!-- Kass&Raftery? Als ja, bedenk wel dat dat om equaility restricties gaat (bij mijn weten). Je kan het natuurlijk gebruiken, maar plaats dan wel de kanttekening, maar is een uitgangspunt natuurlijk.  -->
<!-- Of vertel over betting odds... -->
<!-- Maar het blijft een dichitome beslissing zo.... -->
<!-- Laat je het gemiddelde PBF met range ofzo zien.... zo niet, dan zou dat toevoegen. Dan zie je meer dan een dichitome beslissing. -->
<!-- En ms kan je ook ene keer 1 en 2 gebruiken, gewoon om te zien wat er gebeurt? -->
<!-- Rebecca: -->
<!-- Now, you will say that selecting > .1 with rho .1 is false, but it is not false. It is on the border, so Hi is true.  -->
<!-- In practice it may be true half of the time, because of sampling, but still. -->
<!-- This should be reflected by BF is approx 1. -->
<!-- In your simulation, you use PBF > 3, so this may be hard to reach when the truth is on the border, but that is due to the subjective choice of 3... -->
<!-- So, I already thought that using a smaller population rho value would be of added value, but now I believe you should really include this. -->
<!-- I do see that this would end up in multiple results (once using .1 and 0.2, once (say) 0 or 0.05 and .2; but also when using other cut-off values for PBF). -->
The informative hypothesis, set to be equal across studies,
was $H_i: \rho > .1$.
To examine the performance of the different evidence synthesis methods in a range of scenarios,
several design factors were manipulated.
First was the presence or absence of a true population effect.
Given the informative hypothesis of $H_i: \rho > .1$,
the presence of a true population effect was defined as $\rho = .2$ and
a null effect was defined as $\rho = .1$.
The second design factor was the number of observations per sample $n \in (50, 200, 500, 800)$.
These values were chosen because they correspond to a statistical
<!-- pwr.r.test(c(50, 200, 500, 800), .1) -->
power to reject a false null hypothesis of $\beta \in (.10, .30, .60, .80)$ power, respectively,
assuming $\alpha = .05$ and a known effect size of $\rho = .1$ [@cohenStatisticalPowerAnalysis1988].
Third, we manipulated the number of independent samples (or: replication studies), $k \in (2, 3, 10)$.
Fourth, the reliability of the two correlated variables was varied between $\alpha \in (0.6, 0.8, 1.0)$ to range from questionable to perfect reliability [@nunnallyPsychometricTheory2017].
Questionable reliability is the lowest level considered to be acceptable in social scientific research, and perfect reliability is what is assumed when analyzing correlations between observed items or scale scores.
For all unique combinations of these design factors,
the simulation was repeated 1000 times.

@@

As a benchmark for comparison, we included several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
The first was *vote counting*: counting the number of significant effects.
Although this method is still in use for aggregating conceptual replications,
it is considered bad practice.
Three disadvantages are that vote counting disregards sample size,
reduces statistical power,
and does not quantify the strength of the evidence [@hedgesVotecountingMethodsResearch1980].
Our vote counting algorithm summed the number of one-sided z-tests of a null hypothesis corresponding to the informative hypothesis,
so $H_0: \rho = .1$ and $H_a: \rho > .1$, which corresponds to $H_i$.
The decision criterion was that the hypothesis was supported in the majority of samples.
Thus, for example, if $H_0$ was rejected in three out of five samples, our vote counting algorithm would find overall support for $H_a$ (and, by extension, $H_i$).

@@

For each algorithm, inferential decisions made using the criteria described above were compared to the population status of the hypothesis (true or false).
The resulting confusion matrix gives the number of decisions that were true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN).
These quantities were summarized as sensitivity, $\frac{TP}{TP + FN}$,
the ability to detect an effect given that it was indeed true in the population,
<!-- i think sensitivity is the ability to detect an effect, given that the effect is truly there. I found that its formula is usually presented as TP/(TP+FN)-->
<!--CJ: This is all the same, see https://en.wikipedia.org/wiki/Sensitivity_and_specificity -->
and specificity, $\frac{TN}{TN+FP}$,
the ability to correctly conclude that the informative hypothesis is not supported,
given that it was indeed false in the population. <!-- and specificity is the ability to not acknowledge an effect, given that there indeed is none-->
The overall performance was captured by the accuracy, which represents the total proportion of correct (true positive and true negative) decisions, $\frac{TP + TN}{TP+TN+FP+FN}$.

@@

```{r tabconf, echo = F, message=F}
tab <- read.csv("confusion.csv", stringsAsFactors = FALSE)
names(tab)[1] <- "Metric"
tab$Metric <- gsub("^.+_", "", tab$Metric)
tab <- tab[!tab$Metric == "lr", !grepl("iu", names(tab))]
names(tab) <- gsub("_ic", "", names(tab))
# names(tab)[match(names(rename), names(tab))] <- rename
tab <- tab[tab$Metric %in% c("sensitivity", "specificity", "accuracy"),]
rownames(tab) <- NULL
knitr::kable(tab, digits = 2, caption = "Marginal confusion matrix metrics.")
```
We examined overall model performance across conditions.
PBF had a higher overall accuracy than other algorithms
followed by IPD, then RMA, and finally VC,
see Table \@ref(tab:tabconf).
This higher accuracy was primarily driven by PBF's greater sensitivity to detect a true effect compared to other algorithms.
However, PBF had a lower specificity compared to all other algorithms.
This suggests that the PBF trades a loss of specificity for increased sensitivity.

@@

Figure \@ref(fig:figk) indicates that, for PBF at higher levels of $k$,
lower sensitivity was traded for greater specificity.
The other algorithms did not show this pattern, as their specificity was at a ceiling.
This difference in pattern of effects explains why number of samples had a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
Only VC showed decreasing sensitivity with an increasing number of groups;
this is because the probability of obtaining false negatives increases with the number of groups [as also noted by @hedgesVotecountingMethodsResearch1980].

@@

A third limitation is that the evidence synthesis methods compared here represent different approaches to inference and answer different research questions.
<!-- Their inferential properties are thus not directly comparable. -->
<!-- PBF assesses whether an informative hypothesis is supported across all replication studies; -->
<!-- RMA estimates an average effect size from summary statistics; -->
<!-- IPD estimates an effect size from individual data, accounting for clustering within studies; -->
<!-- and VC counts the number of times a hypothesis was supported. -->
Since each of these methods is optimized for a different purpose,
the present study should not be considered as a comprehensive assessment of their strengths and weaknesses.
<!-- For example, PBF takes a Bayesian approach to inference, and a PBF of 10 in our study means that the (subjective) probability that our informative hypothesis is true is ten times larger than the probability that it is not true. -->
<!-- By contrast, IPD takes a frequentist approach to inference, -->
<!-- and a p-value of $p = .02$ means that there is a 2% probability of observing an effect at least as large as the one found, if the null-hypothesis were true. -->
We nonetheless compare them because of their similar usage in evidence synthesis.
It is up to individual researchers to choose an appropriate method,
guided by the research question and the available information.
For instance, when raw data is unavailable, IPD cannot be used,
and when parameter estimates or effect sizes are not reported, only VC can be used.

Aside from the aforementioned fact that the PBF answers a different research question than the other algorithms,
it is worth noting limitations of the interpretation of the PBF.
The PBF renders support for one specific informative hypothesis versus its complement.
If the informative hypothesis is supported, this does not necessarily mean that it is also true.
Consider the hypothetical example that the informative hypothesis that the earth is flat was supported with $BF = 3.01$.
Although the data support this hypothesis over its complement, the hypothesis is clearly wrong (the earth is spherical).
If we would have evaluated another hypothesis, e.g., the earth is shaped like an American football, it would have received much more support, e.g. $BF = 1000$, even though it is also wrong.
A high Bayes factor thus does not mean that the hypothesis is true.
Conversely, a low Bayes factor merely indicates that the informative hypothesis is not supported,
and does not provide information about the true state of affairs.
A related limitation is that our simulation study used an arbitrary - albeit conventional - threshold for inference [@jeffreys1998theory].
In applied research, it is more sensible to evaluate the weight of evidence, rather than resorting to a rule of thumb.
<!-- In sum, this simulation study suggest that PBF may be a good choice when sensitivity is of utmost importance, but researchers should be mindful of the trade-off between specificity and sensitivity. -->
<!-- The results also highlight the importance of considering sample size, number of samples, -->
<!-- and reliability when designing studies and conducting meta-analyses. Future research should examine the performance of additional algorithms and the impact of other design factors on algorithm performance. -->
<!-- The results indicated that, compared to the other algorithms, -->
<!-- PBF had a higher overall accuracy, -->
<!-- which could be attributed to PBF's relatively greater sensitivity to detect a true effect. -->
<!-- The greater sensitivity of PBF came at a loss of specificity compared to other algorithms. -->
<!-- Specifically, all other algorithms showed ceiling effects for specificity and low base rates for sensitivity. -->

@@

This tutorial demonstrates how to synthesize evidence for an informative hypothesis across heterogeneous replications using the Product Bayes Factor (PBF). 
We assume that users have installed the free open source statistical programming language R [@R-base].
The R-package `bain` version `0.2.9` or later is required, which can be installed by running `install.packages("bain")` in the R console.
The data used in this tutorial are included in the `bain` package,
and have been simulated based on the data presented in [@vanleeuwen2022morality].
A more detailed description of the datasets is found in [@vanleeuwen2022morality];
additionally, the dataset documentation is accessed by running `?synthetic_us`, `?synthetic_dk` or `?synthetic_nl` in the R console.
Van Leeuwen and colleagues conducted a theory-driven, preregistered study to address the research question whether political orientation and moral dispositions are associated.
Suitable data were collected in three countries: the United states of America, Denmark, and the Netherlands.
Each sample contained multiple measures of political orientation and moral dispositions.
In the original publication, the PBF was used to aggregate evidence across scales and countries to obtain an overall measure of support for the central hypothesis.
This tutorial follows the same rationale, but uses only one effect size per sample,
and varies the way this effect size is computed to illustrate the more typical use case where the same informative hypothesis has been studied in different ways in multiple studies.
We will examine the informative hypothesis that self-reported importance of family morality is positively associated with a conservative socio-political orientation.
We load the `bain` library and assign the data to three objects with convenient names:

```{r eval = TRUE, echo = TRUE}
library(bain)
NL <- synthetic_nl
DK <- synthetic_dk
US <- synthetic_us
```

@@

<!-- CJ: Is dit niet teveel herhaling van wat we in de inleiding zeggen?? -->
<!-- In conventional Null-hypothesis significance testing (NHST), as the name implies, data is analysed to generate evidence for or against the null-hypothesis.  -->
<!-- This method evaluates a single hypothesis, the null-hypothesis, and its output is dichotomous in nature: whether a theoretical relationship is probable or improbable to exist given the data.  -->
<!-- However, there are many more hypotheses that can theoretically be tested and researchers may not be interested in the null-hypothesis. They may also not only be interested in whether the hypothesis is probable, but also by how much.  -->
<!-- In the Bayesian framework, the Bayes Factor (BF) is a test statistic that gives the amount of support for any possible desired hypothesis.  -->
<!-- Because the focus of this tutorial is on BES,  -->
We briefly introduce the basic use of the `bain()` function, and how to interpret its output.
We must estimate a model suitable for evaluating our informative hypothesis.
Because both scales consist of multiple items, we can use structural equation modeling (SEM) to perform latent variable regression [see @vanlissaTeacherCornerEvaluating2020]:
<!--EB: For simplicities sake, I for now create simple average scores without taking into account reverse coding. This should likely be improved with factor scores -->
<!-- CJ: Keep the code SUPER simple. No functions etc -->

@@

# Test the hypothesis that the effect size labeled 'beta' is positive
# The argument standardize = TRUE uses the standardized model parameters
bf_nl <- bain(results_nl, hypothesis = "beta > .1", standardize = TRUE)
```

The informative hypothesis in this tutorial is $H_i: \beta > .1$, where $\beta$ (beta) is the standardized regression coefficient.
Instead of a conventional null hypothesis, $H_0: \beta = 0$,
the value of $.1$ was used as a minimal effect size of interest.
The code below illustrates how to obtain a Bayes factor for this informative hypothesis,
using the output of the SEM analysis above.
We can refer to the parameter `beta` by name because we labeled it in the `lavaan` syntax;
if we had not done so, we could find the names of all model parameters by running `get_estimates(results_nl, standardize = TRUE)`.
The results indicate that the hypothesis is supported when compared to its complement.
<!-- , $BF_c `report(bf_nl$fit$BF.c[1])`$. -->
For a more in-depth tutorial on `bain()`, see @hoijtink2019tutorial, and for further guidance on the use of `bain()` for SEM, see @vanlissaTeacherCornerEvaluating2020.

```{r echo = TRUE, eval = TRUE, results = resultaction}
# Test that the effect labeled 'beta' is positive
bf_nl <- bain(results_nl, hypothesis = "beta > .1", standardize = TRUE)
bf_nl
```
